---
title: "Exploration of outliers through Interceptor"
author: "Marlin"
date: "3/7/2022"
params:
  BaseDir: "Z:/"
output: 
  bookdown::html_document2: 
    fig_width: 8
    fig_height: 4
    code_folding: hide
editor_options: 
  chunk_output_type: inline
---


```{r set up markdown settings, echo=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE
)
```


```{r Start enviroment, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(lmtest)
library(lubridate)
library(limma)
library(tidyr)
library(plotly)
library(gridExtra)
library(data.table)
library(formattable)
library(ggpubr)

#Data Files and prep work
source("../../lib/DataProccess.R")
source("../../lib/NormFuncs.R")
source("../../lib/OutlierDetectionFuncs.R")
```

```{r helpful data manipulations, include = FALSE}

SecondAxisFormat <- list(#Second Axis components
  tickfont = list(size=11.7),
  titlefont=list(size=14.6),
  overlaying = "y",
  nticks = 4,
  dtick = 5e5,
  side = "right",
  title = "N1 (GC/L)",
  exponentformat = "e"
)

```



```{r DF Set Up, echo=FALSE}
BaseDir <- params$BaseDir#get the root of the directory where the data is stored
#All the Madison data is contained in these two files#MMSD_Cases_processed
MadisonCaseFN  <-  paste0(BaseDir, "COVID-19_WastewaterAnalysis/data/processed/MMSD_Interceptor_Cases_2_7_22.csv")
LIMSFN <- paste0(BaseDir, "COVID-19_WastewaterAnalysis/data/processed/LIMSWasteData_01-25-21_01-05-22.csv")
#LIMSFN <- paste0(BaseDir, "COVID-19_WastewaterAnalysis/data/processed/LIMSWasteData_02-09-22.csv")

#Importing the Madison case data
LatCaseDF <- ParseData(MadisonCaseFN)%>% 
  mutate(SevenDayMACases = rollapply(data = FirstConfirmed, width = 7, FUN = mean, 
                            na.rm = TRUE,fill=NA))%>%
  filter(Date>ymd("2020-9-10"))%>%
  select(Date, Site, FirstConfirmed,SevenDayMACases)
SiteData <- unique(LatCaseDF$Site)
#Importing the Madison waste water data
LIMSFullDF <- ParseData(LIMSFN)%>%
  filter(Site %in% SiteData)%>%
  mutate(N1Pure = ifelse(is.na(N1),N1,N1),
         N1Pure = ifelse(is.na(N1),N1,N1),
         GeoMeanN12 = exp((log(N1Pure)+log(N1Pure))/2))%>%
  select(Date, Site, N1, BCoV , N1 , PMMoV,
         GeoMeanN12,FlowRate,temperature,equiv_sewage_amt)

#joining the two data frames together
FullDF <- full_join(LatCaseDF,LIMSFullDF, by = c("Date","Site"))

#unique(LIMSFullDF$Site)
```



```{r flowstuff DataPrep}

MadDF <- filter(FullDF,Site=="Madison")%>%
    mutate(FlagedOutliers = IdentifyOutliers(N1,Bin = 21, Action = "Flag"),
           #Manual flagging that method misses due to boundary effect with binning
           FlagedOutliers = ifelse(Date == mdy("01/26/2021") | Date == mdy("01/27/2021"),
                                   TRUE, FlagedOutliers),
           NoOutlierVar = ifelse(FlagedOutliers, NA, N1))

OutlierDF <- MadDF%>%
  filter(FlagedOutliers)
FullDFMassRatio <- FullDF%>%
  filter(!is.na(N1))%>%
  mutate(SC2_mass = (3.785*1e6)*FlowRate*N1)


MissingIntercepter <- FullDFMassRatio%>%
  filter(Site!="Madison",!is.na(FlowRate)|!is.na(N1))%>%
  group_by(Date)%>%
  summarize(N = n())%>%
  filter(N!=5)


FullDFMassRatio.Mad <- FullDFMassRatio%>%
  filter(!is.na(SC2_mass))%>%
  filter(Site=="Madison")
FullDFMassRatio.Inter <- FullDFMassRatio%>%
  filter(!is.na(SC2_mass))%>%
  filter(Site!="Madison")#TempErrorMetric

TempErrorMetric <- FullDFMassRatio.Inter%>%
  group_by(Date)%>%
  summarise(InterSum = sum(SC2_mass),InterSumFlow = sum(FlowRate))%>%
  inner_join(FullDFMassRatio.Mad)%>%
  mutate(MassRatio = SC2_mass/InterSum,
         ErrorRatio = 2*(SC2_mass-InterSum)/(SC2_mass+InterSum),
         MassRatioFlow = FlowRate/InterSumFlow,
         ErrorRatioFlow = 2*(FlowRate-InterSumFlow)/(FlowRate+InterSumFlow))


FullDFMassRatio.Mad <- FullDFMassRatio%>%
  filter(!is.na(SC2_mass))%>%
  filter(Site=="Madison")%>%
  inner_join(FullDFMassRatio.Inter["Date"])
FullDFMassRatio.Inter <- FullDFMassRatio%>%
  filter(Site!="Madison")

```


```{r Stat Look,include=FALSE}
TempErrorMetric%>%
  summarise(MeanError = mean(ErrorRatio, na.rm=TRUE),
            MeanRatio = mean(MassRatio, na.rm=TRUE),
            MedianError = median(ErrorRatio, na.rm=TRUE),
            MedianRatio = median(MassRatio, na.rm=TRUE),
            n())

ToMergeDF <- TempErrorMetric[c("Date","ErrorRatio","MassRatio","MassRatioFlow","ErrorRatioFlow")]

MadErrorRatioDF <- full_join(ToMergeDF,MadDF)%>%
  filter(!is.na(N1),!is.na(MassRatio),!is.na(MassRatio))
MadErrorRatioFilledDF <- MadErrorRatioDF%>%
  setorder(Date)%>%
  fill(ErrorRatio,MassRatio,.direction ="down")
  


MadErrorRatioDF%>%
  filter(!is.na(ErrorRatio))%>%
  group_by(FlagedOutliers)%>%
  summarise(MeanError = mean(ErrorRatio, na.rm=TRUE),
            MeanRatio = mean(MassRatio, na.rm=TRUE),
            MedianError = median(ErrorRatio, na.rm=TRUE),
            MedianRatio = median(MassRatio, na.rm=TRUE),
            n())
MadErrorRatioFilledDF%>%
  filter(!is.na(ErrorRatio))%>%
  group_by(FlagedOutliers)%>%
  summarise(MeanError = mean(ErrorRatio, na.rm=TRUE),
            MeanRatio = mean(MassRatio, na.rm=TRUE),
            MedianError = median(ErrorRatio, na.rm=TRUE),
            MedianRatio = median(MassRatio, na.rm=TRUE),
            n())


HighRatioVec <- ToMergeDF%>%
  filter(MassRatio>1.5)



MadErrorRatioFilledDF%>%
  filter(FlagedOutliers)

MadErrorRatioDF%>%
  ggplot()+
  aes(x=MassRatio,y=N1,color = FlagedOutliers)+
  geom_point(size=5,alpha=.5)+
  scale_y_log10()+
  scale_x_log10()+
  ggtitle("FlagedOutlier vs N1")

MadErrorRatioFilledDF%>%
  ggplot()+
  aes(x=MassRatio,y=N1,color = FlagedOutliers)+
  geom_point(size=5,alpha=.5)+
  scale_y_log10()+
  scale_x_log10()+
  ggtitle("FlagedOutlier vs N1 Filled Data")


t.test(ErrorRatio ~ FlagedOutliers, data = MadErrorRatioFilledDF)



#MadErrorRatioFilledDF
#MadErrorRatioDF
MadErrorRatioDF%>%
  ggplot()+
  aes(x=MassRatioFlow,y=MassRatio,color=FlagedOutliers)+
  geom_point(size=5,alpha=.5)+
  ggtitle("FlowRatio vs MassFlow")


MadErrorRatioDF%>%
  filter(MassRatioFlow==1)
```


```{r Outlier}
KeyOulierPoints <- c(ymd("2021-06-08"),
                     ymd("2021-10-17"),
                     ymd("2021-05-02"),
                     ymd("2021-01-26"))

NonOuliersDF <- MadDF%>%
  mutate(Outlier = ifelse(FlagedOutliers,N1,NA))%>%
           mutate(N1 = NoOutlierVar)%>%
            filter(!(is.na(N1)))

OutLierPlotDF <- MadDF%>%
  mutate(Outlier = ifelse(FlagedOutliers,N1,NA))%>%
           mutate(N1 = NoOutlierVar)%>%
  filter(!(is.na(N1)&is.na(Outlier)))%>%
  ggplot(aes(x=Date))+#Data depends on time 
  geom_line(aes(y=N1,
                color="N1", 
                info = N1),data=NonOuliersDF)+#compares Var to Cases
  geom_point(aes(y=Outlier,
                 color="N1 Outlier",
                 info = Outlier))+
  theme_light() +
  scale_color_manual(values=c("#F8766D","#999999"))

ggplotly(OutLierPlotDF)
#"2021-06-08","2021-10-17","2021-05-02","2021-01-26"
```
Sum of interceptor flows agrees with composite flow except for dates with missing flow information for some interceptors. Purple boxes beneath data are to communicate the Dates with interceptor data partially missing.
```{r Flow total}
MissingFlowData <- FullDFMassRatio.Inter%>%
  group_by(Date)%>%
  summarise(n=n())%>%
  filter(n!=5)%>%
  mutate(MissingSite=TRUE)


FlowPlot <- FullDFMassRatio.Inter%>%
  full_join(MissingFlowData)%>%
  mutate(MissingSite=ifelse(is.na(MissingSite),FALSE,MissingSite))%>%
  filter(!is.na(FlowRate))%>%
  ggplot()+
  geom_col(aes(x=Date,y=FlowRate,fill=Site,shape=MissingSite),position="stack", stat="identity", width = 3) + 
  geom_col(aes(x=Date,y=-.5),data=MissingFlowData,color="Purple",fill="Purple", width = 3)+
  scale_fill_brewer(palette="Spectral") +
  scale_color_manual(values=c("#d60000", "#ffa600", "#7ed05c", "#5c9c43", "#3a56d6")) +
  theme_light() +
  geom_point(data=FullDFMassRatio.Mad,aes(x=Date,y=FlowRate),size=2)+
  ylab("Average daily flow (MGD)") +
  geom_hline(yintercept=median(FullDFMassRatio.Mad$FlowRate), linetype = "dashed", color="black") +
  ggtitle("Average daily flow (MMSD)")

ggplotly(FlowPlot)%>%
  add_annotations( text="Site, Full Data For Day", xref="paper", yref="paper",
                  x=1.02, xanchor="left",
                  y=0.8, yanchor="bottom",    # Same y as legend below
                  legendtitle=TRUE, showarrow=FALSE ) %>%
  layout( legend=list(y=0.8, yanchor="top" ) )
```

```{r MassRatio}
MassBalencePlot <- FullDFMassRatio.Inter%>%
  full_join(MissingFlowData)%>%
  mutate(MissingSite=ifelse(is.na(MissingSite),FALSE,MissingSite))%>%
  #filter(!is.na(FlowRate))%>%
  #filter(!MissingSite)%>%
  ggplot()+
  geom_col(aes(x=Date,y=SC2_mass,fill=Site,shape=MissingSite),position="stack", stat="identity", width = 3) + 
  geom_col(aes(x=Date,y=-3e12),data=MissingFlowData,color="Purple",fill="Purple", width = 3)+
  scale_fill_brewer(palette="Spectral") +
  scale_color_manual(values=c("#d60000", "#ffa600", "#7ed05c", "#5c9c43", "#3a56d6")) +
  theme_light() +
  geom_point(data=FullDFMassRatio.Mad,aes(x=Date,y=SC2_mass),size=1)+
  ylab("N1 gene copies per day") +
  
  ggtitle("SARS-CoV-2, 24h Mass Loading (Madison)")
  

ggplotly(MassBalencePlot)
#4-15
#N1/N1 travel to find more outliers
#Present puzzles to WSLH
#distribution of residuals
#if normalized resid should be iid
```

```{r Cases,include = FALSE}
CasesPlot2 <- FullDFMassRatio.Inter%>%
  filter(!is.na(FlowRate))%>%
  ggplot()+
  geom_col(aes(x=Date,y=FirstConfirmed,fill=Site),position="stack", stat="identity", width = 3) + 
  scale_fill_brewer(palette="Spectral") +
  scale_color_manual(values=c("#29ffff", "#0059ff", "#812fa3", "#a363bc", "#c5a929")) +
  theme_light() +
  geom_point(data=FullDFMassRatio.Mad,aes(x=Date,y=FirstConfirmed),size=1)+
  ylab("N1/N1 gene copies per day") +
  ggtitle("Cases, 24h Mass Loading (Madison)")

ggplotly(CasesPlot2)

unique(inner_join(FullDFMassRatio.Inter,OutlierDF["Date"],by=c("Date"))$Date)
```


```{r,include=FALSE,eval=FALSE}
MassBalencePlot

SC2.mass


FlowPlot

flow


AllOuliersDates <- OutlierDF[["Date"]]
OutlierDatesIntercepters <- unique(inner_join(FullDFMassRatio.Inter,OutlierDF["Date"],by=c("Date"))$Date)




FullDFMassRatio.Inter$Site <- plyr::revalue(FullDFMassRatio.Inter$Site, c("Madison" = "MC",
                                        "MMSD-P11" = "P11",
                                        "MMSD-P18" = "P18",
                                        "MMSD-P2" = "P2",
                                        "MMSD-P7" = "P7",
                                        "MMSD-P8" = "P8"))
FullDFMassRatio.Inter2 <- FullDFMassRatio.Inter%>%
  rename(sewershed=Site,date=Date)%>%
  select(date,sewershed,SC2_mass)#%>%



data.MAD.plot%>%
  mutate(date=mdy(date))%>%
  filter(variable %like% "_SC2_Mass")%>%
  select(date,value,variable,sewershed)%>%
  inner_join(FullDFMassRatio.Inter2)


inner_join(FullDFMassRatio.Inter,OutlierDF["Date"],by=c("Date"))

FullDFMassRatio.Inter%>%
  filter(Site=="MMSD-P8")%>%
  arrange(N1)
#Explore intercepter ratio
```



arranged plots with lines to communicate where flagged outliers are. Red Lines show Where there was a flagged on a day that also has collected interceptor data.
```{r SC2MassRatio, fig.height=6,dpi=500}
StartRange <- mdy("5-01-2021")
EndRange <- mdy("01-30-2022")

AllOuliersDates <- OutlierDF[["Date"]]
OutlierDatesIntercepters <- unique(inner_join(FullDFMassRatio.Inter,OutlierDF["Date"],by=c("Date"))$Date)

a <- OutLierPlotDF+
  scale_x_date(limits = c(StartRange,EndRange))+
  geom_vline(xintercept = AllOuliersDates)+
  geom_vline(xintercept = OutlierDatesIntercepters,color="Red")+
  #geom_vline(xintercept = MissingIntercepter$Date,color="Red")+
  theme(title = element_blank(),
        axis.text.x = element_blank(), 
        axis.title.x = element_blank(),
        legend.position="none")
b <- MassBalencePlot+
  scale_x_date(limits = c(StartRange,EndRange))+
  geom_vline(xintercept = (AllOuliersDates))+
  geom_vline(xintercept = (OutlierDatesIntercepters),color="Red")+
  #geom_vline(xintercept = MissingIntercepter$Date,color="Red")+
  theme(title = element_blank(),
    axis.text.x = element_blank(), 
        axis.title.x = element_blank(),
    legend.position="none")
c <- FlowPlot+
  scale_x_date(limits = c(StartRange,EndRange))+
  geom_vline(xintercept = AllOuliersDates)+
  geom_vline(xintercept = OutlierDatesIntercepters,color="Red")+
  theme(title = element_blank(),
        legend.position="none")



ggarrange(a,b,c,nrow=3,align ="v",legend.grob=get_legend(MassBalencePlot),legend="right")

#Cause of missing data
OutlierDatesIntercepters
```


