---
title: "<center> <h1>Sample degradation analysis</h1> </center>"
author: "<center> <h1>Marlin Lee and Steve Goldstein</h1> </center>"
date: "<center> <h1>August 3, 2021</h1> </center>"
output:
  pdf_document: default
  html_document: default
---
```{r, echo=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

```{r Start enviroment, message=FALSE, warning=FALSE, echo=FALSE}
library(ggpubr)
library(tidyverse)
library(forecast)
library(lmtest)
library(lubridate)
library(lme4)
library(lmerTest)
#Data Files and prep work
source("../Scripts/GenPlotMaking.R")
source("../Scripts/WasteWaterDataProccess.R")
source("../Scripts/CassesDataProccess.R")
source("../Scripts/HelperFunctions.R")
```

```{r DF Set Up, echo=FALSE}
PathStarter="Z:/"

SigWasteFN <- paste0(PathStarter,
      "COVID-19_WastewaterAnalysis/data/raw/TimeSeriesDataforStats072321.csv")

SigDecayDF <- read.csv(SigWasteFN)%>%
  rename(collectionDate = Collection.Date,
         FilteredDate = Filtered.Date)%>%
  mutate(collectionDate = mdy(collectionDate), 
         FilteredDate = mdy(FilteredDate),
         TimePassed = FilteredDate-collectionDate,
         N1.CT = as.numeric(N1.CT))

```




Analysis of TimeSeriesDataforStats072321.csv from 7/23/2021 email from Kayley Janssen  

>One figure I am hoping to include is a time series of filtering to show decay rate or when/if a
significant drop off in signal of SARS-CoV-2 and control viruses in wastewater held at 4 degrees
occurs. We were provided with enough influent from 3 different wastewater treatment plants
where we filtered duplicates of the sample almost every day over a 2 week period and then ran
triplicates(technical replicates) of each filter on qPCR for N1, N2, BCoV, and PMMoV.

Fitting a linear model to log(concentration N1) seems appropriate: the slope of the regression
line estimates the exponential rate of decay of concentration N1. Figure 1 shows the results of
applying a mixed-effects model, estimating a common slope for the three WWTPs and letting
concentration N1 vary between the three. The p-value for the slope coefficient is not significant.  

To investigate why the data don’t fit this model, we fit linear models to each filter replicate
independently. Figure 2 suggests that Wisconsin Rapids Filter replicate A might be an outlier.
Excluding this sample from the dataset and fitting the mixed-effects model yields an estimated
decay rate of -0.041 ± 0.016 GC/(L*day), ( p < 0.05), as shown in Figure 3.  

```{r ploting Funcs, warning=FALSE, message=FALSE}
#

MixedEffectModelSum <- function(DF,Var,Log=TRUE){
  if(Log){
    depenVar <- paste0("log(",Var,")")
  }else{
    depenVar <- Var
  }
  MixedEffectModel <- lmer(as.formula(paste0(depenVar," ~ TimePassed +   (1|WWTP)")),
                         data = DF)
  summary(MixedEffectModel)
}

MixedEffectModelPlot <- function(DF,Var,Title,subtitle,
                                 xlab="Filtered Date",
                                 ylab="N1 (GC/L)",
                                 Log=TRUE){
  if(Log){
    Var2=paste0("log(",Var,")")
  }else{
    Var2=Var
  }
  MixedEffectModel <- lmer(as.formula(paste0(Var2," ~ TimePassed + (1|WWTP)")),
                         data = DF)

  DF3 <- DF%>%
    filter(!is.na(!!sym(Var)))
  
  Pvel <- signif(summary(MixedEffectModel)$coefficients[10],4)
                                    
  
  if(Log){
    DF3$Pred = exp(predict(MixedEffectModel))
  }else{
    DF3$Pred = predict(MixedEffectModel)
  }
  
  FinalGraph <- DF3%>%
    mutate(Filter.Replicate=substr(Filter.Replicate,4,6))%>%
     ggplot()+
     geom_jitter(aes(x=FilteredDate,y=!!sym(Var),
                     color=Filter.Replicate,
                     alpha=as.character(Pvel)),
                 width = .2)+
     geom_line(aes(x=FilteredDate,y=Pred))+
     facet_wrap(~WWTP,nrow = 3)+
    labs(y=ylab,
         x=xlab,
         color = "Filter Replicate",
         title=Title,
         subtitle=subtitle)+
    scale_alpha_manual(values=c(1),labels = c(Pvel),name=expression(p[slope]),guide = 'legend')
  if (Log){
    FinalGraph <- FinalGraph+
      scale_y_log10()
  }
  return(FinalGraph)
}

regressions_data <- function(Var,DF){
  DF%>%
    #mutate(Filter.Replicate=substr(Filter.Replicate,4,6))%>%
    group_by(WWTP)%>%
    summarise(PVal=signif(summary(lm(!!sym(Var)~TimePassed))$coefficients[8], digits = 3))
}

#ggplot latex? subscripts p_slope
DecayPlot <- function(Var,DF,Locate=c(18729,14,25,20),Log=TRUE){
  GraphPlot <- DF%>%
    mutate(Filter.Replicate=substr(Filter.Replicate,4,6))%>%
    ggplot()+
    aes(x=FilteredDate,y=!!sym(Var))+#,color=Filter.Replicate)+
    geom_label(inherit.aes=FALSE, 
          aes(x = as.Date(Locate[1]-Locate[2]*abs(as.numeric(as.factor(WWTP))-2)),
           y = Locate[3],
           label=paste("p[slope]","==",  PVal)),
           size=4,
           show.legend = FALSE,
           parse = TRUE,
           data=regressions_data(Var,DF))+
    geom_jitter(aes(color=Filter.Replicate),width = .2)+
    geom_smooth(method = "lm", se=FALSE)+

    labs(x="Filtered Date",y=Var,color="Filter Replicate")+
    facet_wrap(~WWTP,nrow=3)
  if(Log){
    GraphPlot <- GraphPlot+
      scale_y_log10()
  }
  return(GraphPlot)
}
```


```{r ploting N1, warning=FALSE, message=FALSE,fig.height=4}
SigDecayDF2 <- SigDecayDF%>%
  mutate(N12=N1.GC.L+N2.GC.L)


MixedEffectModelPlot(SigDecayDF, Var = "N1.GC.L",
                     Title="",subtitle="",
                     xlab="Figure 1. Mixed-effects model.",
                     ylab="N1 (GC/L)")


DecayPlot("N1.GC.L",SigDecayDF2)+
  labs(x="Figure 2. Six independent linear regressions.",
       y="N1 (GC/L)")


SigDecayDF2 <- SigDecayDF%>%
  filter(!(WWTP=="WI Rapids"))#&Filter.Replicate=="Rep A"))
SigDecayDF3 <- SigDecayDF%>%
  filter(!(WWTP=="Madison"&Filter.Replicate=="Rep A"))

MixedEffectModelPlot(SigDecayDF2,Var = "N1.GC.L",
                     Title="",
                     subtitle="",
                     xlab="Figure 3. Mixed-effects model excluding WI Rapids filter replicate A",
                     ylab="N1 (GC/L)")


#overlay PMMoV vs N1
#compare slope of different models and SE
#looking at delay in measuring as a way to control LIMS signal
#what model used comes down to formulating what the goal is

#Brian: early august - students coming back - want to be done asap - where we are report useful in that place
#should we devote resources for collection
#make simplest thing to show
#back apply to remove error
#might explain issues
#see relationship between time and difference between testing and collection

#Focus on removing heuristics
#rigorous analysis

#include N1+N2
#N2 are messy with delay
#constant picture between N1 and Pmmov-show that

#2)code review isc conversation
#by early monday lets have something that tells story
#lets have something put together by monday
library(knitr)
library(printr)
```






```{r ploting N2, include=FALSE, warning=FALSE, message=FALSE,fig.height=6}
#This process was repeated with N2 with no significance.

MixedEffectModelPlot(SigDecayDF, Var = "N2.GC.L",
                     Title="",subtitle="",
                     xlab="Figure 1. Mixed-effects model.",
                     ylab="N2 (GC/L)")


DecayPlot("N2.GC.L",SigDecayDF)+
  labs(x="Figure 2. Six independent linear regressions.",
       y="N2 (GC/L)")


MixedEffectModelPlot(SigDecayDF2,Var = "N2.GC.L",
                     Title="",
                     subtitle="",
                     xlab="Figure 3. Mixed-effects model excluding WI Rapids filter replicate A",
                     ylab="N2 (GC/L)")
```




```{r ploting PMMoV, include=FALSE, warning=FALSE, message=FALSE,fig.height=6}

# When the mixed model is fitted for PMMoV we find a much stronger picture indicating an exponential decay relationship.

MixedEffectModelPlot(SigDecayDF, Var = "PMMoV.GC.L",
                     Title="",subtitle="",
                     xlab="Figure 1. Mixed-effects model.",
                     ylab="PMMoV (GC/L)")


DecayPlot("PMMoV.GC.L",SigDecayDF,
          Locate=c(18729,14,5000,2000))+
  labs(x="Figure 2. Six independent linear regressions.",
       y="PMMoV (GC/L)")


MixedEffectModelPlot(SigDecayDF2,Var = "PMMoV.GC.L",
                     Title="",
                     subtitle="",
                     xlab="Figure 3. Mixed-effects model excluding WI Rapids filter replicate A",
                     ylab="PMMoV (GC/L)")
```

```{r ploting BCoV, include=FALSE, warning=FALSE, message=FALSE,fig.height=6}
MixedEffectModelPlot(SigDecayDF, Var = "BCoV.percent.recovery",
                     Title="",subtitle="",
                     xlab="Figure 1. Mixed-effects model.",
                     ylab="Percent bovine coronavirus recovery",
                     Log=FALSE)


DecayPlot("BCoV.percent.recovery",SigDecayDF,
                     Log=FALSE,
          Locate=c(18729,14,10,5))+
  labs(x="Figure 2. Six independent linear regressions.",
       y="Percent bovine coronavirus recovery")


MixedEffectModelPlot(SigDecayDF3,Var = "BCoV.percent.recovery",
                     Title="",
                     subtitle="",
                     xlab="Figure 3. Mixed-effects model excluding WI Rapids filter replicate A",
                     ylab="Percent bovine coronavirus recovery",
                     Log=FALSE)
```


\newpage

This process was repeated with each covariate and summarized in the table one below

```{r}

`N1 using six replicates` <- MixedEffectModelSum(SigDecayDF,"N1.GC.L")[[10]][2,c(1,2,5)]
`N1 excluding WI Rapids rep A` <- MixedEffectModelSum(SigDecayDF2,"N1.GC.L")[[10]][2,c(1,2,5)]
`N2 using six replicates` <- MixedEffectModelSum(SigDecayDF,"N2.GC.L")[[10]][2,c(1,2,5)]
`N2 excluding WI Rapids rep A` <- MixedEffectModelSum(SigDecayDF2,"N2.GC.L")[[10]][2,c(1,2,5)]
`PMMoV using six replicates` <- MixedEffectModelSum(SigDecayDF,"PMMoV.GC.L")[[10]][2,c(1,2,5)]
`PMMoV excluding WI Rapids rep A` <- MixedEffectModelSum(SigDecayDF2,"PMMoV.GC.L")[[10]][2,c(1,2,5)]

`BCoV using six replicates` <- MixedEffectModelSum(SigDecayDF,"BCoV.percent.recovery",Log=FALSE)[[10]][2,c(1,2,5)]
`BCoV excluding Madison rep A` <- MixedEffectModelSum(SigDecayDF2,"BCoV.percent.recovery",Log=FALSE)[[10]][2,c(1,2,5)]

FinalTable <- signif(rbind(`N1 using six replicates`,
                          `N1 excluding WI Rapids rep A`,
                          `N2 using six replicates`,
                          `N2 excluding WI Rapids rep A`,
                          `PMMoV using six replicates`,
                          `PMMoV excluding WI Rapids rep A`,
                          `BCoV using six replicates`,
                          `BCoV excluding Madison rep A`
                          ),2)

#put log in math notation
#Var in 
#sub table
#include slope in graph
#log linear model estimating degergation rate
#Note madison weirdness with star or dag and describe in footnote
#Mark BCoV as non log
#exlude mad because only 2 non 0 data points

colnames(FinalTable) <- c("Decay Rate","SE","P")

knitr::kable(FinalTable,
caption ="Summary of models")
```
