---
title: "WasteWater Analysis 2.0"
author: "Steve Goldstein and Marlin Lee"
output: slidy_presentation
---


```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE,warning = FALSE)
knitr::opts_chunk$set(fig.width=8, fig.height=5) 
library(ggpubr)
library(dplyr)
library(limma)

#Data Files and prep work
source("../Scripts/GenPlotMaking.R")
source("../Scripts/WasteWaterDataProccess.R")
source("../Scripts/CassesDataProccess.R")
source("../Scripts/HelperFunctions.R")


source("../Scripts/WasteShiny/PrepData.R", local = TRUE)

```



```{r start up, include=FALSE}

HFGPop=read.csv("../../UntrackedData/HFGPop.csv.txt")%>%
  select(Site=wwtp_name,Population=pop_served)


#Preset for graphing
ConfigOption=list(
      myseed=1234567890,
      XAxisLabSiz=10,
      YAxisLabSiz=15,
      GenFontSiz=20,
      alphaPoint=.7,
      PointSize=2,
      alphaWeek=.2)
    
```

```{r, include=FALSE}
DayRolling=7


LatCaseDFRoll=RollPerPos(LatCaseDF,"Cases","Tests",Facet="Site",n=DayRolling)%>%
  select(Date,Site,Cases,Per_pos)#%>%
  #mutate(Date=Date-DayRolling/2)

CasePlotMaking = function(SiteS,Method="Cases"){
  SiteCases=LatCaseDF%>%
    filter(Site==SiteS)

  SiteCasesRoll=LatCaseDFRoll%>%
    filter(Site==SiteS)
  
  LIMSSiteDF=LIMSFullDF%>%
   filter(Site==SiteS)


  Date_lim_long=c(min(LIMSSiteDF$Date),max(LIMSSiteDF$Date))
  
  CasePlot <- Buildplot_gen(
      Method,
      MainDF=SiteCases,
      Loc="Site",
      LineDF=SiteCasesRoll,
      Xfreq="30 days",
      DateLimits=Date_lim_long,
      Standards=ConfigOption,
      AxisPos="bottom",
      LineColor="red",
      Colplot=TRUE) #+ theme(ConfigOption)
  return(CasePlot)
}

CasePlotMaking("MMSD-P2")

# LIMSSiteDF=LatCaseDF%>%
#   filter(Site=="Madison")
# Date_lim_long=c(min(LIMSSiteDF$Date),max(LIMSSiteDF$Date))
# aba=LatCaseDF%>%
#   filter(Date %in% seq(Date_lim_long[1]+300,Date_lim_long[2], by = "day"))%>%
#   pull(Cases)
# max(aba,na.rm=T)
```

```{r Loess Model, include=FALSE}
#Gen Both Cases and per Pos plot
LoessSmoothPlot <- function(SiteS,Data=LIMSFullDF,Span=.3,min=1e-5,max=1e5,BoxWidth=.5,Independent="N1",Dependent="N1Error",HasNo=F, Title=T,FullLim=F,SiteLab=T,NoLabel=F){
  PlaceHolder=""
  if(HasNo){
    PlaceHolder="no"
  }
  LIMSDF=Data%>%
    filter(Site==SiteS)%>%
    mutate(Indy=!!sym(Independent),
           Dep=!!sym(Dependent))%>%
    filter(!is.na(Dep),
           !is.na(Indy),
           Dep>0,
           Indy>0)
    
  DFSize=dim(LIMSDF)[1]
  ConWeightModel=loessFit(y=LIMSDF$Indy,
                       x=LIMSDF$Date,
                       weights=rep(1,DFSize),
                       span=Span,
                       min.weight=min,
                       max.weight=max,
                       iterations=10)

  SEWeightModel=loessFit(y=LIMSDF$Indy,
                       x=LIMSDF$Date,
                       weights=LIMSDF$Indy/LIMSDF$Dep,
                       span=Span,min.weight=min,
                       max.weight=max,
                       iterations=10)
  LIMSDF$Pred1 <- ConWeightModel$fitted
  LIMSDF$PredSE <- SEWeightModel$fitted
  LoessPlot=LIMSDF%>%
    mutate(ErrorMin=Indy-Dep,ErrorMax=Indy+Dep)%>%
        mutate(ErrorMin=ifelse(ErrorMin>0,ErrorMin,0))%>%
        ggplot()+
        aes(x=Date)+
        geom_rect(aes(ymin=ErrorMin,ymax=ErrorMax,xmin=Date-BoxWidth,xmax=Date+BoxWidth),color="black",alpha=.25)+
        geom_rect(aes(ymin=Indy,ymax=Indy,xmin=Date-BoxWidth,xmax=Date+BoxWidth),color="black",alpha=.25)+
        geom_line(aes(y=Pred1,color="1"),size=1.25)+
        geom_line(aes(y=PredSE,color=paste0(Independent,"/",Dependent)),size=1.25)
  if(SiteLab){
    LoessPlot=LoessPlot+facet_wrap(~Site)
  }
  LoessPlot=LoessPlot+
        labs(color="Weights Used")+
    ylab(paste0(Independent," (GC/L)"))+
    #Header_theme(ConfigOption)+
      theme(axis.text = element_text(size = ConfigOption$XAxisLabSiz, colour = "black"),
            axis.title = element_text(size = ConfigOption$GenFontSiz, colour = "black"),
            plot.margin = unit(c(0,0,0,0), "cm"))+ 
    theme(legend.position = "bottom")+
    ggtitle("")
  if(Title){
    LoessPlot=LoessPlot+
      ggtitle(paste0("Weighted loess smoothing with ",PlaceHolder," truncation 
                   loessFit(",Independent, " ~ Date, span=",Span,")"))
  }
  if(FullLim){
    XVar=LIMSFullDF$Date
    YVar=pull(LIMSFullDF,Independent)
    XLim <- c(min(XVar),max(XVar))
    YLim <- c(min(YVar[YVar!=0]),max(YVar))
    LoessPlot <- LoessPlot+
      scale_x_date(limits = XLim)+
      scale_y_log10(limits = YLim)
  }else{
    LoessPlot <- LoessPlot+
      scale_y_log10()
  }
  if(NoLabel){
    LoessPlot <- LoessPlot + 
      ggtitle("")+
      ylab("")+
      xlab("")
  }
  return(LoessPlot)
}

```

---

Due to the large variations of the N1 concentration we have tried to create a smoothed version of the data set that captures the underlying shifts of the data. Below is our best attempt at smoothing the N1 concentration data. It shows relation to case data given the spread of shedding of N1. This relationship holds for N2 as well.


```{r,  echo=FALSE}

N1Plot <- LoessSmoothPlot(SiteS="Madison",Span=.1,min=0,max=120,Title = F,FullLim=F,SiteLab=F)+
  xlab("")
PerPosPlot <- CasePlotMaking("Madison",Method="Per_pos")
CasePlot=CasePlotMaking("Madison")

ArrangePlot <- ggarrange(CasePlot,N1Plot,nrow=2,common.legend = TRUE ,align ="hv",legend="bottom")

PlotTitle <- text_grob("Weighted loess smoothing compared to cases", color = "black", face = "bold", size = 16)

annotate_figure(ArrangePlot,top = PlotTitle)

# annotate_figure(ggarrange(PerPosPlot,N1Plot,nrow=2,common.legend = TRUE,align ="hv",legend="bottom"),top = text_grob("Weighted loess smoothing compared to percent positive", color = "black", face = "bold", size = 16))


#princepled ways tp do
```

---

To use standard Error measurements in our smoothing we needed to have it be an unbiased estimate given the concentration. A strong correlation between N1 and N1Error means it needs to be controlled for. Using standard error normalized by N1 gave a meaningful measurement of the consistency of the data.


```{r SE Pred N1 Gen, echo=FALSE}
# ***No strong relationship existed between the modified standard error and other predictors #like Percent bovine recovery and PMMoV***


#ToDo: add custom Weighting
#How to deal with -N2Errors
SimpleN1ErrorN1Model <- function(Independent,Dependent,...){
  TempDF <- LIMSFullDF%>%
    mutate(Indy=!!sym(Independent),
           Dep=!!sym(Dependent))%>%
    filter(!is.na(Dep),
           !is.na(Indy),
           Dep>0,
           Indy>0)
  UnWeightedModelFull2 <- lm(Dep~Indy-1,weights=Indy/Dep,data=TempDF)
  
  TempDF%>%
    mutate(FullDataPredict=predict.lm(UnWeightedModelFull2,.))%>%
    ggplot()+aes(x=Indy)+
    geom_point(aes(y=Dep),size=1)+
    geom_smooth(aes(y=Dep),method="lm",se=F,formula=y~0+x)+
    #geom_point(aes(y=FullDataPredict,color="Predicted"),size=1)+
    scale_x_log10()+
    scale_y_log10()+
    ggtitle("lm(N1Error ~ N1 - 1)")+
    ylab(Dependent)+
    xlab(Independent)+
    theme(aspect.ratio=1)
}
#,Weight="INDDep"
SimpleN1ErrorN1Model(Independent="N1",Dependent="N1Error")+
      theme(axis.text = element_text(size = ConfigOption$XAxisLabSiz, colour = "black"),
            axis.title = element_text(size = ConfigOption$GenFontSiz, colour = "black"),
            plot.margin = unit(c(0,0,0,0), "cm"))

#Add aspect=1
#add slope to plot



```

---

Below are the two versions of the smoothed data capturing different scales of its variations. The boxes are the concentration measured plus or minus the standard error. This gives some measure of how weighted the point is given.


```{r SmoothPlot, echo=FALSE,fig.height=4}


LoessSmoothPlot(SiteS="Madison",Span=.1,min=0,max=120)+
      theme(axis.text = element_text(size = ConfigOption$XAxisLabSiz, colour = "black"),
            axis.title = element_text(size = ConfigOption$GenFontSiz, colour = "black"),
            plot.margin = unit(c(0,0,0,0), "cm"))+ theme(plot.title = element_text(size=16))

LoessSmoothPlot(SiteS="Madison",Span=.3,min=0,max=120)+
      theme(axis.text = element_text(size = ConfigOption$XAxisLabSiz, colour = "black"),
            axis.title = element_text(size = ConfigOption$GenFontSiz, colour = "black"),
            plot.margin = unit(c(0,0,0,0), "cm"))+ theme(plot.title = element_text(size=16))



#Text explaining boxes
#Boxes are N1 +-N1Error
#make boxes stand out less


#Sparse Main Plots in 1 Documents
#auxiliary plots in other locations and loess smoothing:
#Different weightings
#weighting
#Residuals
#Weighted loess smoothing with truncation loessFit(N1 ~ Date, span=.3)
```

---

This is method is extended for the interceptors.


```{r, echo=FALSE,include=TRUE,fig.height=18,fig.width=14}

MapList=lapply(c("Madison","MMSD-P2","MMSD-P7","MMSD-P8","MMSD-P11","MMSD-P18"),LoessSmoothPlot,Span=.3,max=20, Title=F,FullLim=T,NoLabel=T)
#LoessSmoothPlot(SiteS="UW-Sellery",Span=.3,max=20)
#LoessSmoothPlot(SiteS="UW-LakeShore",Span=.3,max=20)

SixPlots=ggarrange(plotlist = MapList, nrow=3,ncol=2, common.legend = T,legend="bottom")
annotate_figure(SixPlots,top = text_grob("Weighted loess smoothing with truncation
     loessFit(N1 ~ Date, span=.3)", color = "black", face = "bold", size = 18))
#No date label, No N1 label
```

---


```{r Missing LIMS, warning=FALSE,fig.height=18,echo=FALSE,include=FALSE}
#LIMS data given does not include all the interceptor data. Older data given does not have N1 Error.

# LIMS data given does not include all the interceptor data. Older data given does not have N1 Error.
```

Also extended to UW data

```{r, warning=FALSE,fig.height=8,echo=FALSE}
#LIMSFullDF
N1CaseCompDorms <- function(SiteS){
  #Dec 21 
  #15
  MinConc=min(LIMSFullDF$Date)
  MaxConc=max(LIMSFullDF$Date)
  
  FallRange=c(MinConc,mdy("12/11/2020"))
  SpringRange=c(mdy("1/25/2021"),MaxConc)
  
  
  N1PlotFull=LoessSmoothPlot(SiteS=SiteS,Data=LIMSFullDF,Span=.3,max=20, Title=F,FullLim=F,NoLabel=T,SiteLab=F)+
    xlab("")
    
  N1PlotSpring=N1PlotFull+
    coord_cartesian(xlim=SpringRange)+
    ggtitle("Spring Semester")+
    theme(axis.text.y = element_blank())+
    theme(plot.margin = unit(c(0,-1,0,-1), "cm"))
  
  
  N1PlotFall=N1PlotFull+
    coord_cartesian(xlim=FallRange)+
    ggtitle("Fall Semester")+
    ylab("N1 (GC/L)")+
    theme(plot.margin = unit(c(0,-1,0,0), "cm"))
  
  CasePlotFull=CasePlotMaking(SiteS,Method="Cases")
  
  CasePlotSpring <- CasePlotFull+
    coord_cartesian(xlim=SpringRange)+
    ylab("")+
    theme(axis.text.y = element_blank())+
    theme(plot.margin = unit(c(0,-1,0,-1), "cm"))
  
  CasePlotFall <- CasePlotFull+
    coord_cartesian(xlim=FallRange)+
    ylab("Cases")+
    theme(plot.margin = unit(c(0,-1,0,0), "cm"))
  
ArrangePlot=ggarrange(N1PlotFall,N1PlotSpring,CasePlotFall,CasePlotSpring,
                      common.legend = TRUE,
                      align ="hv", legend="bottom")
TitleText=text_grob(paste("Case concentration comparison for",SiteS),
                    color = "black", face = "bold", size = 14)
  annotate_figure(ArrangePlot,top = TitleText)
}

N1CaseCompDorms(SiteS="UW-Sellery")
```

---

```{r, warning=FALSE,fig.height=8,echo=FALSE}

N1CaseCompDorms(SiteS="UW-LakeShore")

```

```{r, warning=FALSE,fig.height=12,include=FALSE}
#Per_pos
N1CaseComp <- function(SiteS){
  N1Plot=LoessSmoothPlot(SiteS=SiteS,Data=LIMSFullDF,Span=.3,max=20, Title=F,FullLim=F,NoLabel=T)+
  theme()
  
  CasePlot=CasePlotMaking(SiteS,Method="Cases")
ArrangePlot=ggarrange(N1Plot,CasePlot,nrow=2,common.legend = TRUE ,align ="hv",legend="bottom")
TitleText=text_grob("Weighted loess smoothing compared to cases)", color = "black", face = "bold", size = 14)
  annotate_figure(ArrangePlot,top = TitleText)
}


# N1CaseComp("MMSD-P2")
# N1CaseComp("MMSD-P7")
# N1CaseComp("MMSD-P8")
# N1CaseComp("MMSD-P11")
# N1CaseComp("MMSD-P18")
#HFD second half
#Lims older data
```

---

No clear use for PMMoV and percent bovine coronavirus recovery use is still unclear. Below shows that unusual they do not seem to relate to unusual N1 values.


```{r fig.height = 3,echo=FALSE , fig.width = 13}
SiteOptions2=c("Kenosha","Madison", "Sun Prairie","Wausau")


#PMMoV (GC/L)
PMMoVUseDF=HFGFrame%>%
  filter(Site %in% SiteOptions2)%>%
  mutate(`PMMoV Level`=ifelse(PMMoV<1*10^7,"Low","Medium"))%>%
  mutate(`PMMoV Level`=ifelse(PMMoV>1*10^8,"High",`PMMoV Level`))%>%
  mutate(`PMMoV Level`=factor(`PMMoV Level`,levels=c("High","Medium","Low")))

PMMoVUseDF%>%
  ggplot()+aes(x=PMMoV,y=N1,color=`PMMoV Level`)+geom_point()+scale_y_log10()+
  scale_x_log10()+facet_wrap(~Site,nrow=1)+ggtitle("N1 Vs PMMoV")+
  xlab("PMMoV (GC/L)")+ylab("N1 (GC/L)")

PMMoVUseDF%>%
  ggplot()+aes(x=Date,y=N1,color=`PMMoV Level`)+geom_jitter(height=0,width=.1)+
  ylab("N1 (GC/L)")+ggtitle("PMMoV and N1 appear unrelated")+
  scale_y_log10()+facet_wrap(~Site,nrow=1)

```

---


```{r outliers Per_Bcov, fig.height = 3,echo=FALSE , fig.width = 13}
BCoVUseDF=HFGFrame%>%
  filter(Site %in% SiteOptions2)%>%
  filter(N1<2e6)%>%
  mutate(`Percent BCoV Recovery`=ifelse(Pct_BCoV<.5,"Low","Medium"))%>%
  mutate(`Percent BCoV Recovery`=ifelse(Pct_BCoV>10,"High",`Percent BCoV Recovery`))%>%
  mutate(`Percent BCoV Recovery`=factor(`Percent BCoV Recovery`,levels=c("High","Medium","Low")))

BCoVUseDF%>%
  ggplot()+aes(x=Pct_BCoV,y=N1,color=`Percent BCoV Recovery`)+geom_point()+
  xlab("Percent Bcov Recovery")+
  ylab("N1 (GC/L)")+
  #scale_y_log10()+
  #scale_x_log10()+
  facet_wrap(~Site,nrow=1)+
  ggtitle("N1 Vs %Bcov recovery")
#, scales = "free"
#scale_x_log10()+,color=marked

BCoVUseDF%>%
  ggplot()+aes(x=Date,y=N1,color=`Percent BCoV Recovery`)+geom_jitter(height=0,width=.1)+
  ggtitle("Percent BCoV recovery and N1 appear unrelated")+
  scale_y_log10()+
  ylab("N1 (GC/L)")+
  facet_wrap(~Site,nrow=1)

```





```{r, echo=FALSE,include=FALSE}
# PMMoV and percent bovine coronavirus recovery use is still unclear. Data shows very week and inconsistent trend.

PMMoVN1Relation <- function(Sites){
  LIMSFullDF%>%
  filter(Site==Sites)%>%
  ggplot()+
  aes(x=N1)+
  #scale_colour_viridis_b()+
  geom_point(aes(y=PMMoV))+
  scale_y_log10()+
  scale_x_log10()+
  facet_wrap(~Site)
}
BCoVN1Relation <- function(Sites){
  LIMSFullDF%>%
  filter(Site==Sites)%>%
  ggplot()+
  aes(x=N1)+
  #scale_colour_viridis_b()+
  geom_point(aes(y=BCoV))+
  scale_y_log10()+
  scale_x_log10()+
  facet_wrap(~Site)
}

BCPMMN1Relation <- function(Sites){
  P1=BCoVN1Relation(Sites)
  B1=PMMoVN1Relation(Sites)
  FinPlot=ggarrange(P1,B1,nrow=2)
  return(FinPlot)
}

Sites=c("UW-LakeShore","MMSD-P8","MMSD-P2","MMSD-P11","Madison","MMSD-P7","MMSD-P18","Hudson","Kenosha","Wausau","Oshkosh")
  
#unique(LIMSFullDF$Site)

#lapply(Sites,BCPMMN1Relation)

#lapply(Sites,PMMoVN1Relation)
# LIMSFullDF%>%
#   group_by(Site)%>%
#   summarise(n=n(),Min=min(Date),Dur=max(Date)-min(Date),MeanP=mean(Pop,na.rm=TRUE))%>%
#   arrange(desc(MeanP))
# 
# LatWasteDF%>%
#   group_by(Site)%>%
#   summarise(n=n(),Min=min(Date),Dur=max(Date)-min(Date))


```



