---
title: "Example Data Vignette"
author: "Kyllan Wunder"
vignette: >
  %\VignetteIndexEntry{Example Data Vignette}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
output:
  pdf_document: default
  html_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(Warning = FALSE, message = FALSE, echo = FALSE)
options(warn=-1)
```

#Example data
This data is only meant to be used in vignettes or testing. To view the full data use: data(WasteWater_data, package = "DSIWastewater")

This data comes from our wastewater and case data merged together by site and date. Then filtered with by 3 sites; Green Bay, Janesville, and Mauston. Then a new column is added; geo_mean which is the geometric mean of N1 and N2. This gives roughly 500 rows to work with for examples across different sized sites over a large time frame.

```{r load packages}
library(DSIWastewater)
library(dplyr)
library(ggplot2)
```

This dataframe contains all columns dropped from the main wastewater dataframe. It is merged using [https://github.com/AFIDSI/DSIWastewater/blob/main/docs/data/data_disc.md](https://github.com/AFIDSI/DSIWastewater/blob/main/docs/data/data_disc.md)

```{r Example Data head}
data(aux_info, package = "DSIWastewater")
aux_info%>%
  select(sample_id:bcov_rec_rate)%>%
  head()%>%
  knitr::kable()


aux_info%>%
  select(sample_type:n2_num_ntc_amplify)%>%
  head()%>%
  knitr::kable()

aux_info%>%
  select(avg_sars_cov2_conc:analytical_comments)%>%
  head()%>%
  knitr::kable()

aux_info%>%
  select(inhibition_detect:equiv_sewage_amt)%>%
  head()%>%
  knitr::kable()
```

