---
title: "Calculated Info"
author: "Marlin"
date: "2023-06-15"
vignette: >
  %\VignetteIndexEntry{Calculated Info}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This package contains a couple tools to create useful views of the data not by default in the dataframe. These mostly come in the form of the default case and wastewater dataframe augment function.

# Wastewater augmeanting function
The wastewater function does three things. first it replaces Concentrations below the level of detection (LOD) with LOD / 2 which is considered best practice. Second it calculates a geometric average of the two gene concentrations. Finally it gets a flow and population normalized loged version of the data which previous work has shown the best results.

```{r load and show data}
library(DSIWastewater)
data(WasteWater_data, package = "DSIWastewater")
head(buildWasteAnalysisDF(WasteWater_data))
```
# Case augmeanting function
The case function does two things. first it normalizes the data by population. Second it calculates a rolling sum and average of the data. This needs the population data contained in its own data frame.
```{r explain all / most relavent functions}
data(Case_data, package = "DSIWastewater")
data("pop_data", package = "DSIWastewater")#need to fix pop_data
#head(buildCaseAnalysisDF(left_join(Case_data, pop_data, by = "site")))

```

These functions are slightly out of date and not used in many new analysis projects
```{r explain / show use case and link to other vignettes}
#TODO work on comparing the output of these methods 
```