---
title: "High Frequency Case Data"
author: "Marlin Lee"
vignette: >
  %\VignetteIndexEntry{HFG data}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
output:
  pdf_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, Warning = FALSE, message = FALSE, echo = FALSE)
options(warn=-1)
```

## High Frequency Case Data

This vignette covers the traditionally collected case data during the high frequency period. This case data has four different metrics of cases over a period of 55 weeks.

```{r load and show data}
library(DSIWastewater)
library(dplyr)
library(ggplot2)

data(HFGCase_data, package = "DSIWastewater")
```

The data is broken down by site and date. Site is where the sample was collected from a range of 10 locations. For each date and site there are four metrics of case data. These generally agree but have some slight variation. In this data a case number of less then 5 is represented with -999 for privacy.

```{r catagorical variables, warning = FALSE}

HFGCase_data%>%
  filter(!is.na(ReportedCases))%>%
  ggplot(aes(x = date), size = .2)+
  geom_point(aes(y = ReportedCases, color = "Reported"))+
  geom_point(aes(y = EpisodeCases, color = "Episode"))+
  geom_point(aes(y = CollectedCases, color = "Collected"))+
  geom_point(aes(y = ConfirmedCases, color = "Confirmed"))+
  scale_y_log10()+
  facet_wrap(~site, scales = "free_y")+
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(y = "Number of Cases Reported", color = "Case Collection method", title = "Diffrent High Frequency Cases", x = "Date")
```

Below is the first 6 rows of the dataset

```{r explain / show use case and link to other vignettes}
head(HFGCase_data)%>%
  head()%>%
  knitr::kable()

```


We used this data set in some analysis [here](https://github.com/AFIDSI/Covid19-Wastewater-Analysis/tree/main/results/CoVariants)