---
title: "Longitudinal Waste data"
author: "Marlin"
vignette: >
  %\VignetteIndexEntry{Longitudinal Waste Data}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
output:
  pdf_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, Warning = FALSE, message = FALSE)
options(warn=-1)
```

## Longitudinal data source

This vignette covers the wastewater data collected by the Wisconsin State Lab of Hygiene. This wastewater data covers 118 weeks of 1-5 measurements a week from 82 locations. This has 63 columns where 5 are categorical variables, 9 are measurement columns and the remaining are extra info about the measurements.


```{r load and show data}
#add start and end date?
library(DSIWastewater)
library(dplyr)
library(ggplot2)

data(WasteWater_data, package = "DSIWastewater")
```


The data is broken down by site and data. Site is where it was collected from. The primary column's are N1 and N2 which are two different genes of Covid-19. 

```{r explain most relavent columns, warning = FALSE}
#could put in the work to explain all 62 columns but it should be fine for now
WasteWater_data%>%
  filter(site == "Madison")%>%
  ggplot(aes(x = date), size = .3, alpha = .5)+
  geom_point(aes(y = N1, color = "N1"))+
  geom_point(aes(y = N2, color = "N2"))+
  scale_y_log10()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(y = "Gene Concentration", color = "", title = "Log number of N1 and N2", x = "Date")
```

Below is the first 6 rows of the dataset.

```{r explain show use case and link to other vignettes}
Head_DF <- WasteWater_data%>%
  head()

Head_DF%>%
  select(sample_id:flow)%>%
  knitr::kable()

Head_DF%>%
  select(ph:n1_lod)%>%
  knitr::kable()

Head_DF%>%
  select(n1_loq:n2_loq)%>%
  knitr::kable()

```

This data is our most used data that is used in multiple analyses. A couple of examples are below.

[WPHA Poster](https://github.com/AFIDSI/Covid19-Wastewater-Analysis/blob/main/conclusions/WPHA-Poster/WPHA-Poster.pdf)

[SETAC Poster](https://github.com/AFIDSI/Covid19-Wastewater-Analysis/blob/main/conclusions/SETAC-Poster/SETAC-Poster.pdf)