---
title: "Population Data Vignette"
author: "Kyllan Wunder"
vignette: >
  %\VignetteIndexEntry{Population Data Vignette}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
output:
  pdf_document: default
  html_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, Warning = FALSE, message = FALSE)
options(warn=-1)
```

# Population data

Population Data for Wisconsin's Wastewater treatment plants(WWTP). For cities with more than one WWTP (Madison and Milwaukee) the population is for the whole city.

```{r load packages}
library(dplyr)
library(DSIWastewater)
```

First column is the WWTP site and the next is the population

```{r Population data head} 
data(pop_data, package = "DSIWastewater")
pop_data%>%
  head()%>%
  knitr::kable()

```

Sometimes we split sites by population size into 3 groups Small(below 10,000), Medium(between 10,000 and 100,000), and Large(Over 100,000)

```{r Splitting population data by size}
population_split_by_size <- pop_data %>% 
  mutate(popgroup = case_when(pop >= 100000 ~ "Large",
                             pop >= 10000  ~ "Medium",
                             .default = "Small"))
population_split_by_size%>%
  head()%>%
  knitr::kable()
```


To combine cities that have multiple WWTP you can use code similar to this.

```{r Grouping data by city}
Madison_data <- pop_data %>% filter(grepl("Madison|MMSD", site))
Madison_data%>%
  head()%>%
  knitr::kable()
```

