---
title: "Time Series"
author: "Kyllan Wunder"
vignette: >
  %\VignetteIndexEntry{Time Series Analysis Vignette}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
output:
  pdf_document: default
  html_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(Warning = FALSE, message = FALSE, echo = FALSE)
library(DSIWastewater)
library(dplyr)
library(RcppRoll)
library(tidyverse)
```


## Time Series and Offset Analysis

How to find the offset between wastewater and case data. By offsetting the wastewater and case data 1 day at a time and then finding the highest correlation we can find the offset between the wastewater and case data.


```{r load and show data}
data(Case_data, package = "DSIWastewater")
totalcases <- Case_data %>% 
  group_by(date) %>% 
  summarise(sum_case = sum(conf_case)) %>% 
  mutate(conf_case = roll_sum(sum_case,10,align = "right", fill = NA))

data(WasteWater_data, package = "DSIWastewater")

data(Covariants_data, package = "DSIWastewater")

data(pop_data, package = "DSIWastewater")


WasteWater_data <- WasteWater_data%>%
  left_join(pop_data)

data(example_data, package = "DSIWastewater")
```

**Output is not normal as only exmaple data is used**
---
title: "Time Series"
author: "Kyllan Wunder"
vignette: >
  %\VignetteIndexEntry{Time Series Analysis Vignette}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
output:
  pdf_document: default
  html_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(Warning = FALSE, message = FALSE, echo = FALSE)
library(DSIWastewater)
library(dplyr)
library(RcppRoll)
library(tidyverse)
```


## Time Series and Offset Analysis

How to find the offset between wastewater and case data. By offsetting the wastewater and case data 1 day at a time and then finding the highest correlation we can find the offset between the wastewater and case data.


```{r load and show data}
data(Case_data, package = "DSIWastewater")
totalcases <- Case_data %>% 
  group_by(date) %>% 
  summarise(sum_case = sum(conf_case)) %>% 
  mutate(conf_case = roll_sum(sum_case,10,align = "right", fill = NA))

data(WasteWater_data, package = "DSIWastewater")

data(Covariants_data, package = "DSIWastewater")

data(pop_data, package = "DSIWastewater")


WasteWater_data <- WasteWater_data%>%
  left_join(pop_data)

data(example_data, package = "DSIWastewater")
```

**Output is not normal as only exmaple data is used**

One way of finding the offset.

```{r finding offset}
offsetDF <- OffsetDFMaker(10, as.Date("2020-08-01"), as.Date("2023-01-01"), Case_data, WasteWater_data)

OffsetDF_Plot(offsetDF, "All Wisconsin data over all time")
```

Another way to find the offset.

```{r heatmap correlation finding offset}
heatmapcorfunc(Example_data)
heatmapcorfunc(Example_data, 10)
```

Using found offset can try to see how it changes over time and other factors.

```{r using offset}
covarstarts <- c(as.Date("2020-08-17"),
                 as.Date("2021-03-29"),
                 as.Date("2021-07-05"),
                 as.Date("2021-12-20"),
                 as.Date("2022-03-28"),
                 as.Date("2022-05-23"),
                 as.Date("2022-07-04"))
covarends <- c(as.Date("2021-01-18"),
               as.Date("2021-05-24"),
               as.Date("2021-12-06"),
               as.Date("2022-03-14"),
               as.Date("2022-05-09"),
               as.Date("2022-06-06"),
               as.Date("2022-11-07"))
covarnames <- c("Robin1",
                "Alpha.V1",
                "Delta",
                "Omicron.21K",
                "Omicron.21L",
                "Omicron.22C",
                "Omicron.22B")


OffsetHeatmap("kendall_offset", 0, WasteWater_data, Case_data, pop_data,
              covarstarts, covarends, covarnames, "region", TRUE, TRUE)

```


```{r explain / show use case and link to other vignettes}

```

These functions are used in this analysis:

 - [WPHA Poster](https://github.com/UW-Madison-DSI/Covid19-Wastewater-Analysis/blob/main/conclusions/WPHA-Poster/WPHA-Poster.pdf)
 
 - [NWSS-Data-Analysis-Presentation_5-2-2023](https://github.com/UW-Madison-DSI/Covid19-Wastewater-Analysis/blob/main/conclusions/NWSS-Data-Analysis-Presentation_5-2-2023/5-2_DHS-presentation.pdf)
 