---
title: "Example Data Vignette"
author: "Kyllan Wunder"
vignette: >
  %\VignetteIndexEntry{Example Data Vignette}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
output:
  pdf_document: default
  html_document: default
urlcolor: blue
---

#Example data
This data is only meant to be used in vignettes or testing. To view the full data use: data(WasteWater_data, package = "DSIWastewater")

This data comes from our wastewater and case data merged together by site and date. Then filtered with by 3 sites; Green Bay, Janesville, and Mauston. Then a new column is added; geo_mean which is the geometric mean of N1 and N2. This gives roughly 500 rows to work with for examples across different sized sites over a large time frame.

```{r load packages}
library(DSIWastewater)
library(ggplot2)
```
This columns are all the same as the wastewater data. [https://github.com/AFIDSI/DSIWastewater/blob/main/docs/data/data_disc.md](https://github.com/AFIDSI/DSIWastewater/blob/main/docs/data/data_disc.md)
```{r Example Data head}
data(Example_data, package = "DSIWastewater")
head(Example_data)
```

```{r plotting example data}
ggplot(Example_data, aes(x=date, y=geo_mean, color = site)) +
  geom_point()
```

