---
title: "Example Data Vignette"
author: "Kyllan Wunder"
vignette: >
  %\VignetteIndexEntry{Example Data Vignette}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
output:
  pdf_document: default
  html_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(Warning = FALSE, message = FALSE, echo = FALSE)
options(warn=-1)
```

# Example data
This data is only meant to be used in vignettes or testing. To view the full data use: 
```
data(WasteWater_data, package = "DSIWastewater")
```

This data comes from our wastewater and case data merged together by site and date. 
```
merge(Case_data,WasteWater_data, by = c("date","site"))
```
The data is then filtered with by 3 sites; Green Bay, Janesville, and Mauston. 
```
filter(site=="Green Bay" | site =="Janesville"| site == "Mauston")
```
A new column is added: geo_mean which is the geometric mean of N1 and N2. 
```
mutate(geo_mean = exp((log(N1) + log(N2))/2))
```
This gives roughly 500 rows to work with for examples across different sized sites over a large time frame.

```{r load packages}
library(DSIWastewater)
library(ggplot2)
library(dplyr)
```
These columns are all the same as the wastewater data explained here:  [https://github.com/AFIDSI/DSIWastewater/blob/main/docs/data/data_disc.md](https://github.com/AFIDSI/DSIWastewater/blob/main/docs/data/data_disc.md)
```{r Example Data head}
data(Example_data, package = "DSIWastewater")
head(Example_data)%>%
  head()%>%
  knitr::kable()
```
Here is a simple plot of the geometric mean by site.

```{r plotting example data}
ggplot(Example_data, aes(x=date, y=geo_mean, color = site)) +
  geom_point() +
  ylab("Geometric Mean of N1 and N2") +
  xlab("Date")

```

