---
title: "Longitudinal Case data"
author: "Marlin"
date: "2023-06-15"
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Longitudinal Case Data}
  %\usepackage[utf8]{inputenc}
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, Warning = FALSE, message = FALSE)
options(warn=-1)
```

## Case data

This vignette covers the Case data reported by the Wisconsin Department of Health services. This data covers 150 weeks of daily reports from 63 locations. This has 7 columns where 2 are categorical variables and the other 5 are different reports of Covid spread.


```{r load and show data}
library(DSIWastewater)
library(dplyr)
library(ggplot2)

data(Case_data, package = "DSIWastewater")
```

```{r explain most relavent columns}
main_plot <- Case_data%>%
  filter(site == "Madison")%>%
  ggplot(aes(x = date), size = .3)+
  geom_point(aes(y = tests + 1, color = "tests"))+
  geom_point(aes(y = conf_case + 1, color = "confirmed case"))+
  scale_y_log10()

main_plot

```


```{r explain / show use case and link to other vignettes}
head(Case_data)
```



This data comes is our most used data that is used in many analysis. A couple of the most used ones are below.
[WPHA Poster](https://github.com/AFIDSI/Covid19-Wastewater-Analysis/blob/main/conclusions/WPHA-Poster/WPHA-Poster.pdf)
[SETAC Poster](https://github.com/AFIDSI/Covid19-Wastewater-Analysis/blob/main/conclusions/SETAC-Poster/SETAC-Poster.pdf)