---
title: "Longitudinal Case data"
vignette: >
  %\VignetteIndexEntry{Longitudinal Case Data}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
output:
  pdf_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, Warning = FALSE, message = FALSE)
options(warn=-1)
```

## Case data

This vignette covers the Case data reported by the Wisconsin Department of Health services. This data covers 150 weeks of daily reports from 63 locations. This has 7 columns where 2 are categorical variables and the other 5 are different reports of Covid spread.


```{r load and show data, include = FALSE}
library(Covid19Wastewater)
library(dplyr)
library(ggplot2)

data(Case_data, package = "Covid19Wastewater")
```

```{r explain most relavent columns, echo = FALSE}
main_plot <- Case_data%>%
  filter(site == "Madison")%>%
  ggplot(aes(x = date), size = .3)+
  geom_point(aes(y = tests + 1, color = "Tests"))+
  geom_point(aes(y = conf_case + 1, color = "Confirmed Cases"))+
  scale_y_log10()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(y = "Tests / Cases", color = "", title = "Log number of Cases vs Tests", x = "Date")

main_plot

```

Below is the first 6 rows of the dataset.
```{r link to other vignettes, echo = FALSE}
Case_data%>%
  head()%>%
  knitr::kable()
```



This data comes is our most used data that is used in many analysis. A couple of the most used ones are below.

[WPHA Poster](https://github.com/UW-Madison-DSI/Covid19-Wastewater-Analysis/blob/main/conclusions/WPHA-Poster/WPHA-Poster.pdf)

[SETAC Poster](https://github.com/UW-Madison-DSI/Covid19-Wastewater-Analysis/blob/main/conclusions/SETAC-Poster/SETAC-Poster.pdf)